|

### 2020-05-29 paper/Manuscript.Rmd
AB changes

abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools. The 
  goal of this paper is to lower the threshold to adoption. We introduce the Workflow for
  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible. It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies. This workflow is primarily targeted towards
  scholars working in R on a private computer,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data - but it is flexible enough to accommodate many
  other scenarios.
  To facilitate widespread adoption, the
  WORCS principles are implemented in the R-package `worcs`, which offers an
  RStudio 
  project template that is, effectively, a "one-click solution" for R users. 
  This paper introduces the workflow, discusses how it meets different 
  requirements of open science, and provides examples of the implementation of
  `worcs` in R.

authornote: |
  This is a preprint, not a peer-reviewed publication. Comments and
  contributions are welcome. Please contact the lead author.
  All authors of this paper, except the first, are listed in alphabetical order.
  Author contributions are detailed in the document 'authors.csv' on the project
  repository, at [https://github.com/cjvanlissa/worcs](https://github.com/cjvanlissa/worcs).
  The lead author is supported by a NWO Veni grant (NWO grant number
  VI.Veni.191G.090).
  We acknowledge Jeroen Ooms for offering feedback and suggesting the use of the
  R package gert.
keywords          : "open science; reproducibility; r; dynamic document generation; version control, FAIR"
wordcount         : "8257"

@@

Many researchers are motivated to adopt best practices for open
science and enjoy these benefits. And yet, the question of "how" to 
apply open science practices can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by presenting a simple workflow 
that meets most requirements open science, and is based on best practices:
The Workflow for Open Reproducible 
Code in Science (WORCS). WORCS is a lightweight approach to open science and
computational reproducibility for scholars working in R on a private computer,
conducting research that involves academic prose, analysis code,
and (optionally) tabular data - but it is flexible enough to accommodate many
other scenarios. The workflow (Figure \@ref(fig:workflow)) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible. WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them. It can also be used in the absence of top-down support or guidelines
to produce work in accordance with best practices.
Although the principles underlying this workflow are universal, WORCS has first
been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
The R-package `worcs` offers a project template for RStudio
[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
a "one click solution". This paper introduces the WORCS workflow, discusses how
it meets best practices for open science, and illustrates the use of the `worcs`
package.

@@

Although Open Science is advocated by many, it does not have a unitary definition
[for an effort to ground open science in first principles, see 
@tennantValuePropositionOpen2020].
Instead, several guidelines exist for practices that are considered "open".
The TOP-guidelines are one of the most influential concrete operationalisations
of Open Science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis; and 8) replication of published results.
WORCS defines the goals of open science in terms of these guidelines, and is
designed to facilitate meeting each of these guidelines, with one exception: We
do not address replication of published results, because replication relates to
the subject of a research project, not to its execution.
We group the remaining seven guidelines into three categories: citation (1),
sharing (2-5), and preregistration (6-7).

@@

Although there is a slight learning curve to transitioning to DDG, the
investment really pays off. Time saved from painstakingly copy-pasting output
and manually formatting text soon outweighs the investment of switching to a new
program. Moreover, human error in manually copying results is eliminated.
When revisions require major changes to the analyses, all results, figures and
tables can be automatically updated. The flexibility in output formats also 
means that a manuscript can be rendered to presentation format, or even to a
website or blog post. Moreover, the fact that code can be run each time the
document is compiled encourages researchers to work reproducibly,
and allows reviewers and/or readers verify reproducibility simply by re-compiling
the document.
In sum, while writing academic papers in a programming environment might seem
counter-intuitive at first, this approach is much more amenable to the needs of
academics than most word processing software. It prevents mistakes, and saves
time.

In the R-implementation of WORCS, we recommend centering a research project
around one dynamically generated 'RMarkdown' document, which includes all 
analysis code. Longer scripts can be
stored in `.R` files, and called from the main document using the `source()`
function. When a reader or reviewer compiles this document, all code is run
automatically, thus verifying computational replicability. The 'RMarkdown' 
document can be automatically formatted in many styles, including
APA style [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020],
a host of other scientific formats 
[see @allaireRticlesArticleFormats2020], and as plain 'Markdown'.

@@

It is worth noting that,
although this paper focuses on the most widely used 'Git'
cloud storage service, 'GitHub', two notable alternatives exist: 
'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
version 0.1.2, the `worcs` template dialog accepts 'https' addresses from 
repositories hosted on either of these platforms interchangeably.
The package has a
[vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) with
additional instructions for using alternative platforms. In 
order to present the workflow in a platform-agnostic way, we will refer to such
cloud hosting services as "remote repositories" in the remainder of this
paper.
However, `worcs` has been tested most extensively with 'GitHub'.

@@

Many solutions exist to ensure computational reproducibility. These solutions typically work by
enveloping your research project in a distinct "environment" that only has access to programs that
are explicitly installed, and maintaining a record of these programs. These solutions differ in
user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a 'Docker container': An environment that behaves like a
virtual computer, which can be stored like a sort of time capsule, and
identically reinstated on a user computer, or in the cloud.
This is effective, and also preserves software outside the R environment (e.g.,
preprocessing tools for neuroimaging data). An excellent Docker-based workflow
was recently introduced by @peikertReproducibleDataAnalysis2019. This workflow
is based on many of the same principles as WORCS, and the two approaches are
conceptually and practically compatible. However, the workflow described by 
Peikert and Brandmaier can be difficult to set up for novice users. An R-package
is currently in development to facilitate this transition [@repro]. Another 
user-friendly Docker-based workflow is offered by the cloud-based collaboration
platform 'Digital Ocean'. Although Digital Ocean enjoys some popularity among 
scientists, it is not clear whether this platform's
["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/)
is GDPR compliant - it seems to focus more on limiting liability than on
compliance. The necessity to upload data for analysis is a risk with any
cloud-computing solution when human participants are involved.

@@

The R-package `worcs` automates most steps of the workflow
for R users.
Several arguments support the choice to implement this workflow in R.
First, R and all of it extensions are free open source software, which
make it a tool of choice for open science.
Second, all tools required for an open science workflow are implemented in R,
most of these tools are directly accessible through the user interface of RStudio,
and some of them are actively developed by the team behind RStudio.
Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
Finally, R is highly interoperable:
Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
Fortran, and Python [@allaireMarkdownPythonEngine2020]. There are excellent free
resources for learning R, such as the book 'R for Data Science' [@grolemundDataScience2017].

Working with R is simplified immensely by using the RStudio integrated
development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
RStudio automates and streamlines tedious or complicated aspects of working with
the tools used in WORCS, and many are embedded directly into the visual user
interface.
This makes RStudio a comprehensive solution for open science research projects.
Another important feature of RStudio is project management.
A project bundles writing, analyses, data, references, etcetera, into a
self-contained folder, that can be uploaded entirely to a 'Git' remote repository and downloaded
by future users. The `worcs` R-package installs a new RStudio project template.
When a new project is initialized from this template, the bookkeeping required
to set up an open science project is performed automatically.

@@

If data can be shared openly, they can simply be pushed to a public 'Git'
repository, along with the analysis code. This way, others can download the
entire repository and reproduce the analyses from start to finish. 
If data cannot be shared, researchers should aim to safeguard the potential for
computational reproducibility and secondary analysis as much as possible. WORCS
recommends two solutions to accomplish this goal.
The first solution is to publish a *checksum* of the original data file.
Think of a checksum as a 32-character summary, or as a "one word" description,
of the contents of a file.
Any change to the file will result in a different checksum. 
Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditor that running the public analysis code on the private data results in the published results of their work.

@@

Note that these functions can be used to store multiple data files, if necessary.
As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
As explained above, this is recommended, because `.csv` files are human- and
machine readable, and because data typically need to be in tabular format
for analysis. Many types of data can be represented as a table - text
corpora can have one row per document, and EEG- or EKG waveforms can have one
row per measurement occasion. The `prepare_data.R` file should document any
steps required to convert these data into tabular format. 
If this is not an appropriate format for the data, readers are encouraged to follow the
development of the `repro` package [@repro], which will offer greater support for
reproducibly storing and retrieving diverse data formats.

@@

The `worcs` workflow facilitates preregistration,
primarily by explicitly inviting a user to write the preregistration document.
To this end, `worcs` imports several preregistration templates from the R-package
`prereg` [@austPreregMarkdownTemplates2019], including templates from
organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and
from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
When initializing an RStudio project with the `worcs` project template, one of
these preregistration templates can be selected, which will generate a file
called `preregistration.Rmd`. This file should be used to document study plans,
ideally prior to data collection. 
Within the workflow, the 'Git' cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a 'Git' remote repository.
By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
it is distinct from all other commits, and easily findable by people and machines.

@@

Being familiar
with 'Git' remote repositories opens doors to new forms of collaboration:
In the open source
software community, continuous peer review and voluntary collaborative acts by
strangers who are interested in a project are commonplace
[see @adolphOpenBehavioralScience2012].  This kind of collaboration is budding
in scientific software development as well; for example, the lead author of this
paper became a co-author on several R-packages after submitting pull requests
with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018], and two co-authors of this
paper became involved by contributing pull requests to `worcs`. It is also
possible to invite such collaboration by 
[opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues)
for tasks that still need to be accomplished, and tag known collaborators to
address them, or invite external collaborators to contribute their expertise.

@@

WORCS provides a user-friendly and
lightweight workflow for open, reproducible research, that meets all
TOP-guidelines, except replication.
Nevertheless, there are clear directions for future developments.
Firstly, although the workflow is currently implemented only in R, it is 
conceptually relevant for users of other statistical programming languages, most
notably Python [@10.5555/1593511].
We welcome efforts to implement WORCS in other platforms.
Secondly, even when a project is in R, it may have dependencies outside of the
R environment that cannot be managed using `renv`.
It is beyond the scope of the `worcs` package to support the incorporation of
tools outside of R, or to containerize a project so that it can
be identically reinstated on a different system or virtual machine.
Therefore, another
important area of future development is to enable such extensions of the
workflow. To this end, we are collaborating with @repro on the development of
`repro`, an R-package that will facilitate setting up reproducibility tools such
as 'Make' and 'Docker'.
This will make it possible for researchers
to enhance a WORCS project with these strict reproducibility tools,
or vice versa, to initiate a WORCS project within a strictly reproducible
environment generated using `repro`. 
This will combine the strengths of WORCS with those of the workflow proposed
by @peikertReproducibleDataAnalysis2019.

>>>>

abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools. The 
  goal of this paper is to lower the threshold to adoption. We introduce the Workflow for
  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible. It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies. This workflow is primarily targeted towards
  scholars working in R on a private computer,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data - but it is flexible enough to accommodate many
  other scenarios.
  To facilitate widespread adoption, the
  WORCS principles are implemented in the R package `worcs`, which offers an
  RStudio 
  project template that is, effectively, a "one-click solution" for R users. 
  This paper introduces the workflow, discusses how it meets different 
  requirements of open science, and provides examples of the implementation of
  `worcs` in R.

authornote: |
  This is a preprint, not a peer-reviewed publication. Comments and
  contributions are welcome. Please contact the lead author.
  All authors of this paper, except the first, are listed in alphabetical order.
  Author contributions are detailed in the document 'authors.csv' on the project
  repository, at [https://github.com/cjvanlissa/worcs](https://github.com/cjvanlissa/worcs).
  The lead author is supported by a NWO Veni grant (NWO grant number
  VI.Veni.191G.090).
  We acknowledge Jeroen Ooms for offering feedback and suggesting the use of the
  R package gert.
keywords          : "open science; reproducibility; r; dynamic document generation; version control"
wordcount         : "8257"

@@

Many researchers are motivated to adopt current best practices for open
science and enjoy these benefits. And yet, the question of "how" to 
apply open science practices can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by presenting a simple workflow 
that meets most requirements open science, and is based on best practices:
The *Workflow for Open Reproducible Code in Science* (WORCS).
WORCS is a lightweight approach to open science and
computational reproducibility for scholars working in R on a private computer,
conducting research that involves academic prose, analysis code,
and (optionally) tabular data - but it is flexible enough to accommodate many
other scenarios. The workflow (Figure \@ref(fig:workflow)) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project (including data, code, and a report) open and reproducible.
WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them. It can also be used in the absence of top-down support or guidelines
to produce work in accordance with best practices.
Although the principles underlying this workflow are universal, WORCS has first
been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
The R package `worcs` [@worcspackage2020] offers a project template for RStudio
[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
a "one click solution". This paper introduces the WORCS workflow, discusses how
it meets best practices for open science, and illustrates the use of the `worcs`
package.

@@

Although Open Science is advocated by many, it does not have a unitary definition
[for an effort to ground open science in first principles, see 
@tennantValuePropositionOpen2020].
Instead, several guidelines exist for practices that are considered "open".
The TOP-guidelines are one of the most influential operationalisations
of Open Science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis; and 8) replication of published results.
WORCS defines the goals of open science in terms of these guidelines, and is
designed to facilitate meeting each of these guidelines, with one exception: We
do not address replication of published results, because replication relates to
the subject of a research project, not to its organization or execution.
We group the remaining seven guidelines into three categories: citation (1),
sharing (2-5), and preregistration (6-7).

@@

Although there is a slight learning curve to transitioning to DDG, we strongly
believe that the investment will pay off for users of the workflow, as it did for us.
Time saved from painstakingly copy-pasting output
and manually formatting text soon outweighs the investment of switching to a new
program. Moreover, human error in manually copying results is eliminated.
When revisions require major changes to the analyses, all results, figures and
tables can be automatically updated. The flexibility in output formats also 
means that a manuscript can be rendered to presentation format, or even to a
website or blog post. Moreover, the fact that code can be run each time the
document is compiled encourages researchers to work reproducibly,
and allows reviewers and/or readers verify reproducibility simply by re-compiling
the document.
In sum, while writing academic papers in a programming environment might seem
counter-intuitive at first, this approach is much more amenable to the needs of
academics than most word processing software. It prevents mistakes, and saves
time.

In the R-implementation of WORCS, we recommend centering a research project
around one dynamically generated 'RMarkdown' document, which includes all 
analysis code. Longer scripts can be
stored in `.R` files, and called from the main document using the `source()`
function. When a reader or reviewer compiles this document, all code is run
automatically, thus verifying computational replicability. The 'RMarkdown' 
document can be automatically formatted in many styles, including
APA style [thanks to the R package `papaja`, @austPapajaPrepareReproducible2020],
a host of other scientific formats 
[see @allaireRticlesArticleFormats2020], and as plain 'Markdown'.

@@

It is worth noting that,
although this paper focuses on the most widely used 'Git'
remote repository service, 'GitHub', two notable alternatives exist: 
'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org).
In 
order to present the workflow in a platform-agnostic way, we will refer to such
cloud hosting services as "remote repositories" in the remainder of this
paper.
As of
version 0.1.2, `worcs` has been tested with these three remote repository services,
but has been optimized for use with 'GitHub'.
A [vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) is included
with instructions for using alternative platforms. 
When initializing a new project using the `worcs` project template,
a 'https' remote repository address will usually work best with the authentication procedure outlined in the
[setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).

@@

Many solutions exist to ensure computational reproducibility. These solutions typically work by
enveloping your research project in a distinct "environment" that only has access to programs that
are explicitly installed, and maintaining a record of these programs. These solutions differ in
user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a 'Docker container': An environment that behaves like a
virtual computer, which can be stored like a sort of time capsule, and
identically reinstated on a user computer, or in the cloud.
This is effective, and also preserves software outside the R environment (e.g.,
preprocessing tools for neuroimaging data). A Docker-based workflow
was recently introduced by @peikertReproducibleDataAnalysis2019. This workflow
is based on many of the same principles as WORCS, and the two approaches are
conceptually and practically compatible. However, the workflow described by 
Peikert and Brandmaier can be difficult to set up for novice users. 
The accompanying R package `repro`
is currently in development to facilitate this transition [@repro]. Another 
user-friendly Docker-based workflow is offered by the cloud-based collaboration
platform 'Digital Ocean'. Although Digital Ocean enjoys some popularity among 
scientists, it is not clear whether this platform's
["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/)
is GDPR compliant - it seems to focus more on limiting liability than on
compliance. The necessity to upload data for analysis is a risk with any
cloud-computing solution when human participants are involved.

@@

The R package `worcs` automates most steps of the workflow
for R users.
Several arguments support the choice to implement this workflow in R.
First, R and all of it extensions are free open source software, which
make it a tool of choice for open science.
Second, all tools required for an open science workflow are implemented in R,
most of these tools are directly accessible through the user interface of RStudio,
and some of them are actively developed by the team behind RStudio.
Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fourth, R is well-supported;
a vibrant and inclusive online community exists for R,
developing new R methods and packages,
and providing support and tutorials for existing ones.
Finally, R is highly interoperable:
Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
Fortran, and Python [@allaireMarkdownPythonEngine2020]. There are excellent free
resources for learning R, such as the book 'R for Data Science' [@grolemundDataScience2017].

Working with R is simplified immensely by using the RStudio integrated
development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
RStudio automates and streamlines tedious or complicated aspects of working with
the tools used in WORCS, and many are embedded directly into the visual user
interface.
This makes RStudio a comprehensive solution for open science research projects.
Another important feature of RStudio is project management.
A project bundles writing, analyses, data, references, et cetera, into a
self-contained folder, that can be uploaded entirely to a 'Git' remote repository and downloaded
by future users. The `worcs` R package installs a new RStudio project template.
When a new project is initialized from this template, the bookkeeping required
to set up an open science project is performed automatically.

@@

If data can be shared openly, they can simply be pushed to a public 'Git'
repository, along with the analysis code. This way, others can download the
entire repository and reproduce the analyses from start to finish. 
If data cannot be shared, researchers should aim to safeguard the potential for
computational reproducibility and secondary analysis as much as possible. WORCS
recommends two solutions to accomplish this goal.
The first solution is to publish a *checksum* of the original data file [@rivest1992md5].
Think of a checksum as a 32-character summary, or as a "one word" description,
of the contents of a file.
Any change to the file will result in a different checksum^[It is theoretically possible but improbable that random changes to a file will result in the same checksum.]. 
Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditor that running the public analysis code on the private data results in the published results of their work.

@@

Note that these functions can be used to store multiple data files, if necessary.
As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
As explained above, this is recommended, because `.csv` files are human- and
machine readable, and because data typically need to be in tabular format
for analysis. Many types of data can be represented as a table; for example,
text corpora can have one row per document, and EEG- or ECG waveforms can have a
row per measurement occasion and a column per channel.
The `prepare_data.R` file should document any
steps required to convert these data into tabular format. 
If this is not an appropriate format for the data, readers are encouraged to follow the
development of the `repro` package [@repro], which will offer greater support for
reproducibly storing and retrieving diverse data formats.

@@

The `worcs` workflow facilitates preregistration,
primarily by explicitly inviting a user to write the preregistration document.
To this end, `worcs` imports several preregistration templates from the R package
`prereg` [@austPreregMarkdownTemplates2019], including templates from
organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and
from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
When initializing an RStudio project with the `worcs` project template, one of
these preregistration templates can be selected, which will generate a file
called `preregistration.Rmd`. This file should be used to document study plans,
ideally prior to data collection. 
Within the workflow, the 'Git' cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a 'Git' remote repository.
By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
it is distinct from all other commits, and easily findable by people and machines.

@@

Being familiar
with 'Git' remote repositories opens doors to new forms of collaboration:
In the open source
software community, continuous peer review and voluntary collaborative acts by
strangers who are interested in a project are commonplace
[see @adolphOpenBehavioralScience2012].  This kind of collaboration is budding
in scientific software development as well; for example, the lead author of this
paper became a co-author on several R packages after submitting pull requests
with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018], and two co-authors of this
paper became involved by contributing pull requests to `worcs`. It is also
possible to invite such collaboration by 
[opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues)
for tasks that still need to be accomplished, and tag known collaborators to
address them, or invite external collaborators to contribute their expertise.

@@

WORCS provides a user-friendly and
lightweight workflow for open, reproducible research, that meets all
TOP-guidelines, except replication.
Nevertheless, there are clear directions for future developments.
Firstly, although the workflow is currently implemented only in R, it is 
conceptually relevant for users of other statistical programming languages, most
notably Python [@10.5555/1593511].
We welcome efforts to implement WORCS in other platforms.
Secondly, even when a project is in R, it may have dependencies outside of the
R environment that cannot be managed using `renv`.
It is beyond the scope of the `worcs` package to support the incorporation of
tools outside of R, or to containerize a project so that it can
be identically reinstated on a different system or virtual machine.
Therefore, another
important area of future development is to enable such extensions of the
workflow. To this end, we are collaborating with @repro on the development of
`repro`, an R package that will facilitate setting up reproducibility tools such
as 'Make' and 'Docker'.
This will make it possible for researchers
to enhance a WORCS project with these strict reproducibility tools,
or vice versa, to initiate a WORCS project within a strictly reproducible
environment generated using `repro`. 
This will combine the strengths of WORCS with those of the workflow proposed
by @peikertReproducibleDataAnalysis2019.
